Efficient and Scalable Bipartite Matching with Fast Beta Linkage (fabl)

### 2022-09-11 ba_resubmit/response_to_reviewers.tex

	\section*{Editor}

	\pointRaised{E}{%
	The paper has been reviewed by a referee and an associate editor and I have read the manuscript independently, before looking at the reports.  When I confront my notes and overall impression with the reviewers’ comments, I find myself in substantial agreement with the reviewers.

	As indicated by both reviewers, the work expands on previous work by Sadinle, introducing computational advances that are achieved by relaxing some requirements in the original formulation of the problem.  The close connections with the previous work and the extent of the novel contributions are not adequately explained.  This makes it hard for the reader to understand what is new and why the new contributions are valuable.  A careful editing of the presentation is needed to address this shortcoming.}

	\reply{%
\textbf{We thank the Editor, the AE, and the reviewer for their feedback, and pointing out our strengths of the paper and areas that we can improve the quality of our paper. We have revised our paper in the abstract and introduction as to make our contributions more clear in relation to prior work.  Furthermore, we have carefully edited the paper to address issues that were pointed out by the Editorial Board.} \todo
	}

@@

	\pointRaised{E}{%
	Regarding the presentation, the AE points out a number of problems with notation, typos, and various inconsistencies.  Independently, I found many similar problems.  The intersection between what I found and what the AE found is non-empty and so is the symmetric difference.  Also, the AE found some issues that I did not find and vice versa.  So, I am almost certain that an additional review would uncover more problems.  Now, the results presented in the paper seem plausible as do the broad strokes of the derivations, but I must admit that I was not able to follow all the details. This, in large part, was because of the issues that I just mentioned, and I believe that most readers would find it difficult to follow the developments, as these issues are exceedingly distracting.
	}

	\reply{
\textbf{We appreciate these comments, and have done our utmost to address the ones listed by yourself and the AE. Furthermore, we have carefully gone through the paper ourselves, where we have corrected other issues with notation, typos, and various inconsistencies.} \todo
	}

	\pointRaised{E}{%
	There are problems with the notation, which in places is not defined, in others is used before being defined, and in others yet is used inconsistently. For example, $I_{obs()}$ is never defined, $n_{12(Z)}$ first appears on p. 5, but is not defined (in passing) until the bottom of p. 7, and it seems to become D on p. 8.
	}

	\reply{
\textbf{$I_{obs}()$ is now defined. We have defined $n_{12}(Z)$ at its first occurrence (and removed the error with $D$). We have checked for other notational inconsistencies.
	}
}

@@

        	\pointRaised{E}{%
	Formatting of all displayed math must be checked and fixed, especially as far as punctuation is concerned, as commas seem to be missing in multiple places. For example, in Equation (1), there should be a comma after 0 and after 1, and this is by no means an isolated occurrence.
	}

	\reply{
		Punctuation for all math has been checked.
	}

\pointRaised{E}{%
	Confusion can also arise from lack of clarity in the exposition.  Take comment 4 from the AE, for example.  Whose marriage certificates and whose birth records are involved?  What records are in $X_1$ and what records are in $X_2$?  The reader should not be expected to go back to Newcombe’s paper to clarify the issue.
	}

	\reply{%
\textbf{We agree with this comment, and we have removed this example as it does not add to the paper.}
}

@@

	\setcounter{responsectr}{0}

	\pointRaised{R1}{%
	This manuscript addresses the bipartite matching problem in data linkage and proposes a new fast computation version of “Beta Linkage” to allow such matching to take place for pairs of large databases without resorting to ad hoc blocking. It is built upon the Felegi-Sunter model of record linkages, where linkages are made within strata of agreement between records across the two databases, and relies heavily on advances made by Sadinle (2017). Overall, the paper presents a practical implementation of a modified Beta Linkage approach to	matching. It demonstrates that even though the new method suffers from the limitations of not seamlessly enforcing one-to-one matching, it is feasible for large datasets with modest numbers of features on which to match. Though the novelty of the presented methodology is primarily in the computational approach, I believe it to be a real advance for practical implementation of record linkage in modern applications.

	I have no major concerns about the content or accuracy of the presentation. However, substantial improvements are warranted to confirm my impression. In addition to the recommendation made by Reviewer 1 to include more clarity of the novel contributions of this manuscript, I believe that the explanation of the method should be written more clearly, the simulation comparisons should be repeated/expanded, and various features of the simplification should be fleshed out a bit more. Finally, more practical advice would be welcome. All of these points are more specifically described in the itemized list below
	}

	\reply{%
	\textbf{We thank the reviewer (AE) for noticing our computational novelties that expand upon the prior limitations of the prior work of Sadinle (2017). Our revised paper clarifies our novel contributions, improves upon the exposition of the writing, provides practical guidance for users, and expands the simulations. We expand on these points further below.} \todo
	}

@@

%	\textcolor{red}{OPTION 2: We agree that the notation provided conflates the records themselves with the record indices. However, this notation was established in Sadinle 2017 and has been replicated in further work (Wortman 2019, Alishen-Guendel, one of Jerry's students), and we have decided to use it as well for consistency within the literature.}
%	\textcolor{blue}{We appreciate the suggestion, however, we believe it would be confusing to the existing prior work by Fellegi and Sunter (1969) that has been established as well as others that have extended the work, such as Sadinle (2014+), Wortman (2019), and Asheshin-Gundel (2020+). Given that it's widely used in this literature, we continue to utilize it.}
}

\pointRaised{R1}{%
Section 2.2, beta distribution: the $n_{12}(.)$ notation needs to be defined.
}
\reply{%
	\textbf{This has been defined.}
}

\pointRaised{R1}{%
Section 3, 4th full paragraph: My understanding of the example of matching birth records and
marriage certificates (presuming monogamy) may not quite be consistent with the algorithm as
presented, as I would think that one certificate (element in X1) would match multiple birth
records (elements in X2) which violates the size ordering. Can the algorithm be easily adjusted
for this reverse sizing?
}

\reply{
	\textbf{Thank you for the interesting comment/question. In this paper, our main goal and applications deal with bipartite record linkage (such that we can make fair comparison with prior work), which is consistent with the algorithm that we present in the paper. We have removed the confusing text mentioned by the reviewer. We are working on extensions beyond bipartite record linkage in other work, where the reverse sizing is not an issue; however, this is beyond the scope of this paper.}
}

@@

%\textcolor{blue}{This is a good point. I wonder if we could perhaps remove a lot of the notation by Sadinle/FS and redo the notation such that the ordering wouldn't matter. Also, it would make more sense to follow with the examples that we use in the paper -- NLTCS and EL Salvador so that we don't confuse the reader. Let's talk through this point as this would be a big change. If the algorithm and notation can be easily modified to reverse things, this would be a nice addition to the paper and improvement over Sadinle.}

\pointRaised{R1}{%
5. Section 3, full model: $I_obs(.)$ seems to be taken from a different paper dealing with missing fields, which is not at all discussed in this manuscript. This consideration of missing fields and the accompanying discussion of missing data mechanism should be included if such data is to be considered for this manuscript. This comment also applies to the $Gamma^{obs}$ notation in the following subsection.
}

\reply{%
	\textbf{Our missingness assumptions and the definition of $I_{obs}()$ have been added near end of Section 3.} \textcolor{red}{Think about where best to put this.}
}

@@

\pointRaised{R1}{%
	Section 3, full model: The square brackets seem to be misplaced, as the exponent includes indices of the sum contained in the brackets.
}

\reply{%
	\textbf{The expression is now corrected.}
}

\pointRaised{R1}{%
Section 3.1, I found the motivational descriptions in this section to be somewhat confusing, in that they seem to mix the posterior behavior (e.g., parameters are updated though standard multinomial-Dirichlet…) and model construction (as a function of pi, Z is a series of successes…). This subsection might be easier to understand if its purpose was more bluntly described in an opening sentence or paragraph and streamlined to focus only on the full conditional distributions for a Gibbs Sampler. In addition, the mixed motivation may have led to errors in intermediate conclusions. In particular, the derivation of the probability of $\Gamma_{.j}$ seems to be missing a factor of $u_{fl}^I(\gamma_{ij}^f=l)$. This is not important to the eventual full conditional of Z, but does not seem to be correct for the distribution of Gamma as claimed.
}

\reply{%
	We have revised our presentation of the Gibbs sampler with this feedback in mind. The main body of the text now presents the final full conditionals, with revised derivations now in the Appendix. In particular, we have included an extra step in the derivation of $p(\Gamma_{.j})$ that shows exactly how we divide by $\prod_{i = 1}^{n_1}\prod_{f = 1}^{F} \prod_{\ell = 1}^{L_f}  u_{fl}^{I(\gamma_{ij}^f=l)}$ and use proportionality to arrive that the distribution presented.
}

@@

\pointRaised{R1}{%
Section 3.1, first equation on page 8: There is an extra parenthesis in the exponent of (1-$\pi$).
}
\reply{%
	\textbf{This has been corrected.}
}

\pointRaised{R1}{%
Section 3.1, pmf for $Gamma_{.j}$ and full conditional for Z: The notation in the final equation may be more clear if you replace i with $z_j$, as in $w_{z_j j}$. I believe that “$n_1+1$” should be “$n_1+ j$” in several places.
}

\reply{%
	\textbf{We have changed $w_{ij}$ to $w_{z_j, j}$ in the pmf for $\Gamma_{.j}$ and the full conditional for $Z_j$. We have been sure that “$n_1+1$” has been changed to “$n_1+ j$” throughout the paper.}
}

\pointRaised{R1}{%
	Section 3.1, last equations on page 8: There seems to be some lack of specificity in describing these equations as full conditional distributions vs. steps in the Gibbs Sampler algorithm. It would be helpful to provide a brief justification of integrating out pi.
}
\reply{%
	\textbf{We have revised the presentation of the Gibbs sampler based on this feedback. The main body of the paper presents the results, and revised derivations (including justification for integrating out $\pi$) have been placed in the Appendix.}
}

@@

\pointRaised{R1}{%
	Section 4.2, SEI procedure. I’m not sure I follow the description of the SEI procedure. (1) My understanding of SEI is that for patterns with lots of possible pairs, no single pair is likely to (should) be identified in the posterior distribution. (2) Thus, you take a small subsample of the records of size $S < H_{j_p}$, and store these in $R^{SEI}$ rather than the complete index of pairs in each R. (3) Can you quantify the computational savings due to this method? (4) Do you have advice for how small S can be? (5) In your simulations, how does the use of this method or choice of S affect the results?
}

\reply{%
	We have numbered each sentence in the comment above to clearly address each point. (1) and (2) are correct understandings of the SEI procedure. For (3), we quantify computational savings of SEI in Section 4.2, where we state:
	''Rather than storing \(n_1 \times n_2\) record labels, SEI allows us to store at most \(n_2 \times P \times S\) labels, regardless of how large \(n_1\) might be." SEI does not affect the computation time of the Gibbs sampler, since the complexity of the sampler is $O(n_2 P)$, regardless of how many comparison vectors are of pattern $P$.

	For (4), we have included the following sentence in the paper ``In practice, we recommend $S=10$, as this reduces the number of stored indices for highly unlikely record pairings, but does is not likely to eliminate any of the indices for plausible matches."

@@

	For (5), we have included the following sentence in the paper ``Choosing $S$ too low, like $S=1$ or $S=2$, can concentrate undue mass on unlikely matches and distort linkage results." 
}

\pointRaised{R1}{%
Section 4.4, Assumptions and definitions should be included in the statement of this lemma.
}

\reply{%
	\textbf{The lemma now includes assumptions and definitions for all variables used within the statement.}
}

@@

\pointRaised{R1}{%
	Section 5.2 and 5.3 simulations – are these a single simulated datasets? Do these same results hold over multiple simulated datasets, Or was this just a chance result for each setting?
}

\reply{%
	In the caption under Figure 3, we explain that we have 100 pairs datasets for each level of overlap and error. Thus the table represents results from 900 pairs of datasets. We have added this description in the body of the paper as well for clarity. 
}

\pointRaised{R1}{%
Section 5.3 simulation – RR results are not presented.
}

\reply{%
	We have updated the Figure 4 with these results. Since it is presented alongside NPV and PPV, for which values close to 1 indicate strong performance, we have presented decision rate (DR) results, instead of rejection rate (RR) results.
}

### 2022-09-11 ba_resubmit/kundinger_resubmit.tex
revisions to paper

\begin{abstract}
Recently, researchers have developed Bayesian versions of the Fellegi Sunter model for record linkage. These have the crucial advantage of quantifying uncertainty from imperfect linkages.  However, current implementations of Bayesian Fellegi Sunter models are computationally intensive, making them challenging to use on larger-scale record linkage tasks.  We propose a variation on Bayesian Fellegi Sunter models that we call fast beta linkage, or \texttt{fabl}.  Specifically, we propose independent prior distributions over the matching space, allowing us to use hashing techniques that reduce computational overhead. This also allows us to complete pairwise record comparisons over large datasets through parallel computing and to reduce memory costs through a new technique called storage efficient indexing. Through simulations and two case studies, we show that \texttt{fabl} has markedly increased speed with minimal loss of accuracy.
\end{abstract}

@@

In many data analysis tasks, analysts seek to identify duplicate records across two databases. This is an increasingly important task in ``data cleaning,'' and is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others \citep{christen_2012, gutman2013bayesian, DalzellReiter18, tang2020}. In this article, we consider bipartite record linkage, which merges two databases together that contain duplications across but not within the respective databases.

@@

The independent pairwise matching assumption from \cite{fellegi_theory_1969} is popular mainly for its mathematical simplicity, but can be unreasonable in practice. In many situations, we know that there are no duplications within a database, meaning that each record should be linked with at most one other record. Thus, when the procedure results in many to one matches, some of these links must be false. Many extensions to \cite{fellegi_theory_1969} resolve these false matches as a post-processing step \citep{jaro1989}, but this model misspecification can still lead to poor results \citep{sadinle_bayesian_2017}.

Alternatively, one can embed one-to-one matching requirements into the model specification itself \citep{gutman2013bayesian, liseo_2011}, at an additional computational cost. \cite{Larsen05} employed a Metropolis-Hastings algorithm to only allow sampling matches that respected one-to-one assumptions, but such algorithms exhibit slow mixing due to the combinatorial nature of the constrained matching space. \cite{fortunato_2010} used simulated annealing to target the space of matches permitted under the one-to-one constraint, but the method is quite computationally intensive and, to our knowledge, has not been applied on databases with more than 100 records. \cite{sadinle_bayesian_2017} proposed the Beta Record Linkage model (\texttt{BRL}), using a prior over the space of bipartite matchings to strictly enforce one-to-one requirements throughout a Gibbs sampler. Additionally, he introduced a class of loss functions that allows for a flexible estimation of the linkage structure, such that the modeller can weight the relative importance of false positives and false negatives, and identify records pairings to be decided through clerical review. \texttt{BRL} was shown to work on larger tasks than previous one-to-one methods, but in our experience, it becomes slow when applied to files with more than a few thousand records.

In this paper, we propose fast beta linkage (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability. Following the suggestion in \cite{wortman2019}, we relax the one-to-one matching requirement of \texttt{BRL} and propose independent priors over the matching space, creating a ``many-to-one'' model for record linkage. This allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large datasets via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing. We demonstrate that even in cases where a bipartite matching is desired, \texttt{fabl} (with a simple post-processing procedure) provides accurate estimation of the linkage structure and other parameters, more information through which to asses model misspecification, and greatly enhanced speed. Open source software to use \texttt{fabl} in \texttt{R} is available through \href{https://github.com/briankundinger/parlrdev}{Github}.

In what follows, Section \ref{sec:review-of_prior-work} reviews the work of \cite{fellegi_theory_1969} and \cite{sadinle_bayesian_2017}. Section \ref{sec:fast-beta-linkage}, proposes the \texttt{fabl} model, derives the Gibbs sampler for posterior inference, and provides the loss function used to calculate the Bayes estimate for the bipartite matching. Section \ref{sec:efficiency} introduces the hashing technique and storage efficient indexing used to increase the speed of calculatoins and the scale of linkage tasks amenable to \texttt{fabl}. Sections \ref{sec:simulations} and \ref{sec:case-studies} demonstrate the speed and accuracy of \texttt{fabl} through simulation studies and case studies of homicides from the El Salvadoran Civil War and the National Long Term Care Study.

@@

\textcolor{red}{Consider two databases $\bf{X}_1$ and $\bf{X}_2$ containing records $\{x_{1i}\}_{i=1}^{n_1}$ and $\{x_{2j}\}_{j=1}^{n_2}$ respectively. Without loss of generality, denote files such that $n_1 \geq n_2$. We follow the convention established by \cite{sadinle_bayesian_2017} and say ``record i $\in X_1$" rather than the more compact $x_{1i}$ in order to avoid double subscripts.} In the context of bipartite matching, we assume that there are duplications across, but not within, each database. Under this framework, the set of matches across databases can be represented in two equivalent ways. First, we may use a matrix $\Delta \in \{0, 1\}^{n_1 \times n_2}$, where
\begin{align}
	\Delta_{ij} =
	\begin{cases}
		1, & \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
		0, & \text{otherwise}.\\
	\end{cases}
\end{align}
Though intuitive, this sparse matrix representation can become cumbersome for large linkage tasks. More compactly, bipartite matching also can be viewed as a labeling $\bm{Z} = (Z_1, \ldots, Z_{n_2})$ for the records in database $\bm{X}_2$ such that 
\begin{align}
	Z_{j} =
	\begin{cases}
		i, & \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
		n_1 + j, & \text{if records}\;  j \in \bm{X}_2 \; \text{does not have a match in database}\; \bm{X}_1. \\
	\end{cases}
\end{align}
We can go back and forth between the two using $\Delta_{ij} = I(Z_j = i),$ where $I(\cdot) = 1$ when the expression inside the parentheses is true, and $I(\cdot) = 0$ otherwise.

Denote the set of matches by $\bm{M} = \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 1\}.$, and the set of non-matches by 
$\bm{U} =  \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 0\}.$ The record linkage task can be viewed as identifying the sets of  $\bm{M}$ and  $\bm{U}.$ We refer to record pairs that are estimated as matches as ``links'', and record pairs that are estimated as non-matches as ``non-links''.

@@

Intuitively, co-referent records (those that refer to the same entity) should be similar; records that are not co-referent should not be similar. \cite{fellegi_theory_1969} proposed encoding this is using a comparison vector $\bm{\gamma}_{ij}$ computed for each record pair $(i,j)$ in $\bm{X}_1 \times \bm{X}_2.$ Denote the number of criteria for comparing records by $F$, such that $\bm{\gamma}_{ij} = (\gamma_{ij}^1, \gamma_{ij}^2, \ldots, \gamma_{ij}^f, \ldots, \gamma_{ij}^F).$ In most cases, $\bm{\gamma}_{ij}$ consists of one comparison for each feature shared between the two datasets.

The simplest way to compare two records is to check for agreement or disagreement, and this is commonly used for categorical variables. For more complex measurements, we can take into account partial agreement to more richly characterize the comparison; for numerical data, we can use absolute difference, and for text data, we can use string distance metrics such as Levenstein or Jaro-Winkler distance \citep{cohen2003comparison}. We then can set thresholds that allow us to represent comparisons through discrete levels of disagreement \citep{bilenko2006riddle, elmagarmid_duplicate_2007}. Let $\mathcal{S}_f(i,j)$ denote a general similarity measure for feature $f$ of records $i$ and $j,$ where the range of $\mathcal{S}_f$ can be divided into $L_f +1$ intervals denoted by $I_{f0}, I_{f1}, \ldots, I_{fL_f}$. Following convention, $I_{f0}$ represents the highest level of agreement (inclusive of complete agreement) and $I_{fL_f}$ represents the highest level of disagreement (including complete disagreement). Thus, we can construct comparison vectors in the following way: 
$\gamma_{ij}^f = \ell \; \text{if} \; \mathcal{S}_f(i,j) \in I_{f\ell}.$
The choice of $I_{f\ell}$ are application specific, which we discuss in our simulation and case studies.

In practice, it is not feasible to make all-to-all record comparisons as the computational complexity is of order $O(n_1 \times n_2).$ The most common solution is to utilize blocking, which places similar records into partitions, or ``blocks,'' to reduce this computational burden \citep{steorts_comparison_2014, murray2016probabilistic}. In deterministic blocking, the modeller chooses a field thought to be highly reliable, and only compares records that agree on that field. The record linkage method is then applied independently across all blocks, which can be done in parallel for additional speed gains. However, blocking on an unreliable field can lead to missed matches, making this form of blocking undesirable in some situations \citep{steorts_comparison_2014}.

@@

\subsection{Fellegi-Sunter Model}
\label{fellegi-sunter}
The comparison data $\bm{\gamma}_{ij}$ are not sufficient to determine whether a record is a match or non-match because of errors that naturally occur in the data. This motivated \cite{fellegi_theory_1969} to consider the likelihood ratio
\begin{align}
	\label{eqn:wts}
	w_{ij} = \frac{P(\gamma_{ij} \mid \Delta_{ij} = 1)}{P(\gamma_{ij} \mid \Delta_{ij} = 0)}
\end{align}
as a weight to estimate if record pairs are a match or non-match. This ratio will be large when there is strong evidence of the pair being a match, and small otherwise.

In practice, $P(\cdot \mid \Delta_{ij} = 1)$ and $P(\cdot \mid \Delta_{ij} = 0)$ are unknown to the modeler, and thus must be estimated. Under the Fellegi Sunter model, we assume both that comparison vectors are independent given their matching status, and that the match status of the record pairs are independent. This allows us to model the comparison data using mixture models in the following way:
\begin{align}
	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 1 \stackrel{iid}{\sim} \mathcal{M}(\bm{m}), \\
	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 0  \stackrel{iid}{\sim} \mathcal{U}(\bm{u}), \notag\\
	& \Delta_{ij}   \stackrel{iid}{\sim} \text{Bernoulli}(p), \notag
\end{align}
where $p$ is the proportion of records that match, and $\bm{m}$ is the set of parameters for the matching record pairs $m_{fl} = P(\gamma_{ij}^f = l|Z_j =i)$ for all fields $f$ and agreement levels $\ell_f$, and $\bm{u}$ is defined similarly for non-matching pairs. These unknown parameters are often estimated using the EM algorithm \citep{Winkler1988}.

%is the set of parameters $u_{fl} = P(\gamma_{ij}^f = l|Z_j \neq i)$ for the non-matching pairs.

After estimating $\bm{m}$ and $\bm{u}$, the Fellegi Sunter method uses thresholds $T_m$ and $T_u$ such that all record pairs with $w_{ij} > T_m$ are declared as links, and all those with $w_{ij} < T_u$ are declared as non-links (with those such that $T_u < w_{ij} < T_m$ left undetermined and subject to clerical review). This is a problem in practice as transitive closures can be violated. Since each $w_{ij}$ is calculated independently, there can be situations in which both $w_{ij} > T_m$ and $w_{i'j} > T_m$, so both $(i,j)$ and $(i, j')$ are declared as links, even though such a matching is not allowed by assumption.

To address this issue, \cite{jaro1989} proposed solving the following linear sum assignment problem:
\begin{align}
	\label{eqn:jaro}
	&\max_{\Delta} \sum_{i=1}^{n_1} \sum_{j=1}^{n_2} w_{ij} \Delta_{ij} 
	\quad \text{subject to} \quad \Delta_{ij} \in \{0,1\}; \\
	& \quad \quad \sum_{i=1}^{n_1}  \Delta_{ij}  \leq 1, j=1,2, \ldots n_2; \; \text{and} \notag\\
	& \quad \quad\sum_{i=1}^{n_2}  \Delta_{ij}  \leq 1, j=1,2, \ldots n_1. \notag
\end{align}

>>>>

\begin{abstract}
Recently, researchers have developed Bayesian versions of the Fellegi Sunter model for record linkage. These have the crucial advantage of quantifying uncertainty from imperfect linkages.  However, current implementations of Bayesian Fellegi Sunter models are computationally intensive, making them challenging to use on larger-scale record linkage tasks.  We propose a variation on Bayesian Fellegi Sunter models that we call fast beta linkage, or \texttt{fabl}.  Specifically, we propose independent prior distributions over the matching space, allowing us to use hashing techniques that reduce computational overhead. This also allows us to complete pairwise record comparisons over large datasets through parallel computing and to reduce memory costs through a new technique called storage efficient indexing. Through simulations and two case studies, we show that \texttt{fabl} has markedly increased speed with minimal loss of accuracy.

\textcolor{red}{BEKA/JERRY: Reviewer 2 wanted a specific reference to Sadinle in the abstract. I remember we actively decided not to do that. What do you think?}
\end{abstract}

@@

In many data analysis tasks, analysts seek to identify duplicate records across two databases. This is an increasingly important task in ``data cleaning,'' and is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others \citep{christen_2012, gutman2013bayesian, DalzellReiter18, tang2020}. In this article, we consider bipartite record linkage, which merges two databases that contain duplications across but not within the respective databases.

@@

The independent pairwise matching assumption from \cite{fellegi_theory_1969} is popular mainly for its mathematical simplicity, but can be unreasonable in practice. In many situations, we know that there are no duplications within a database, meaning that each record in one file should be linked with at most one other record in the other file. Thus, when the procedure results in many-to-one matches, some of these links must be false. Many extensions to \cite{fellegi_theory_1969} resolve these false matches as a post-processing step \citep{jaro1989}, but this model misspecification can still lead to poor results \citep{sadinle_bayesian_2017}.

Alternatively, one can embed one-to-one matching requirements into the model specification itself \citep{gutman2013bayesian, liseo_2011}, at an additional computational cost. \cite{Larsen05} employed a Metropolis-Hastings algorithm to only allow sampling matches that respected one-to-one assumptions, but such algorithms exhibit slow mixing due to the combinatorial nature of the constrained matching space. \cite{fortunato_2010} used simulated annealing to target the space of matches permitted under the one-to-one constraint, but the method is computationally intensive and, to our knowledge, has not been applied on databases with more than 100 records. \cite{sadinle_bayesian_2017} proposed the Beta Record Linkage model (\texttt{BRL}), using a prior distribution over the space of bipartite matchings to strictly enforce one-to-one requirements throughout a Gibbs sampler. Additionally, he introduced a class of loss functions that allows for a flexible estimation of the linkage structure, such that the modeler can weight the relative importance of false positives and false negatives, and identify records pairings to be decided through clerical review. \texttt{BRL} was shown to work on larger tasks than previous one-to-one methods, but in our experience, it becomes slow when applied to files with more than a few thousand records.

In this paper, we propose fast beta linkage (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability. Following the suggestion in \cite{wortman2019}, we relax the one-to-one matching requirement of \texttt{BRL} and propose independent priors over the matching space. This allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large datasets via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing. We demonstrate \texttt{fabl} provides accurate estimation of the linkage structure and greatly enhanced speed. Open source software to use \texttt{fabl} in \texttt{R} is available through \href{https://github.com/briankundinger/parlrdev}{Github}.

In what follows, Section \ref{sec:review-of_prior-work} reviews the work of \cite{fellegi_theory_1969} and \cite{sadinle_bayesian_2017}. Section \ref{sec:fast-beta-linkage}, proposes the \texttt{fabl} model, provides the Gibbs sampler for posterior inference, and shows the loss function used to calculate the Bayes estimate for the bipartite matching. Section \ref{sec:efficiency} introduces the hashing technique and storage efficient indexing used to increase the speed of calculations and the scale of linkage tasks amenable to \texttt{fabl}. Sections \ref{sec:simulations} and \ref{sec:case-studies} demonstrate the speed and accuracy of \texttt{fabl} through simulation studies and case studies of homicides from the El Salvadoran Civil War and the National Long Term Care Study.

@@

\textcolor{red}{Consider two databases $\bm{X}_1$ and $\bm{X}_2$ containing records $\{x_{1i}\}_{i=1}^{n_1}$ and $\{x_{2j}\}_{j=1}^{n_2}$ respectively. Without loss of generality, denote files such that $n_1 \geq n_2$. We follow the convention established by \cite{sadinle_bayesian_2017} and say ``record i $\in X_1$" rather than the more compact $x_{1i}$ in order to avoid double subscripts.} In the context of bipartite matching, we assume that there are duplications across, but not within, each database. Under this framework, the set of matches across databases can be represented in two equivalent ways. First, we may use a matrix $\Delta \in \{0, 1\}^{n_1 \times n_2}$, where
\begin{align}
	\Delta_{ij} =
	\begin{cases}
		1, & \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
		0, & \text{otherwise}.\\
	\end{cases}
\end{align}
Though intuitive, this sparse matrix representation can become cumbersome for large linkage tasks. More compactly, bipartite matching also can be viewed as a labeling $\bm{Z} = (Z_1, \ldots, Z_{n_2})$ for the records in database $\bm{X}_2$ such that 
\begin{align}
	Z_{j} =
	\begin{cases}
		i, & \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
		n_1 + j, & \text{if records}\;  j \in \bm{X}_2 \; \text{does not have a match in database}\; \bm{X}_1. \\
	\end{cases}
\end{align}
We can go back and forth between the two using $\Delta_{ij} = I(Z_j = i),$ where $I(\cdot) = 1$ when the expression inside the parentheses is true, and $I(\cdot) = 0$ otherwise.

Denote the set of matches by $\bm{M} = \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 1\}$, and the set of non-matches by 
$\bm{U} =  \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 0\}.$ The record linkage task can be viewed as identifying the sets of  $\bm{M}$ and  $\bm{U}.$ We refer to record pairs that are estimated as matches as ``links'', and record pairs that are estimated as non-matches as ``non-links''.

@@

Intuitively, matching records (those that refer to the same entity) should be similar; records that are not non-matching should not be similar. \cite{fellegi_theory_1969} proposed encoding this is using a comparison vector $\bm{\gamma}_{ij}$ computed for each record pair $(i,j)$ in $\bm{X}_1 \times \bm{X}_2.$ Denote the number of criteria for comparing records by $F$, such that $\bm{\gamma}_{ij} = (\gamma_{ij}^1, \gamma_{ij}^2, \ldots, \gamma_{ij}^f, \ldots, \gamma_{ij}^F).$ In most cases, $\bm{\gamma}_{ij}$ consists of one comparison for each feature shared between the two datasets. We refer to set of all comparison vectors $\bm{\gamma}_{ij}$ as $\Gamma$.

The simplest way to compare two records is to check for agreement or disagreement, and this is commonly used for categorical variables. For more complex measurements, we can take into account partial agreement to more richly characterize the comparison; for numerical data, we can use absolute difference, and for text data, we can use string distance metrics such as Levenstein or Jaro-Winkler distance \citep{cohen2003comparison}. We then can set thresholds that allow us to represent comparisons through discrete levels of disagreement \citep{bilenko2006riddle, elmagarmid_duplicate_2007}. Let $\mathcal{S}_f(i,j)$ denote a general similarity measure for feature $f$ of records $i$ and $j,$ where the range of $\mathcal{S}_f$ can be divided into $L_f +1$ intervals denoted by $I_{f0}, I_{f1}, \ldots, I_{fL_f}$. Following convention, $I_{f0}$ represents the highest level of agreement (inclusive of complete agreement) and $I_{fL_f}$ represents the highest level of disagreement (including complete disagreement). Thus, we can construct comparison vectors in the following way: $\gamma_{ij}^f = l \; \text{if} \; \mathcal{S}_f(i,j) \in I_{fl}.$ The choice of $I_{fl}$ are application specific, which we discuss in our simulation and case studies.

\textcolor{red}{(I don't love this right here, but I don't really know where else to put it. But if it won't cause a problem with the reviewers, I'll just leave it.)}In practice, it is not feasible to make all-to-all record comparisons as the computational complexity is of order $O(n_1 \times n_2).$ The most common solution is to utilize blocking, which places similar records into partitions, or ``blocks,'' to reduce this computational burden \citep{steorts_comparison_2014, murray2016probabilistic}. In deterministic blocking, the modeler chooses a field thought to be highly reliable, and only compares records that agree on that field. The record linkage method is then applied independently across all blocks, which can be done in parallel for additional speed gains. However, blocking on an unreliable field can lead to missed matches, making this form of blocking undesirable in some situations \citep{steorts_comparison_2014}.

@@

\subsection{Fellegi-Sunter Model}
\label{fellegi-sunter}

Under the Fellegi Sunter model, we assume first, that comparison vectors are independent given their matching status, and second, that the match status of the record pairs are independent. This allows us to model the comparison data using mixture models in the following way:
\begin{align}
	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 1 \stackrel{iid}{\sim} \text{Multinomial}(\bm{m}), \\
	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 0  \stackrel{iid}{\sim} \text{Multinomial}(\bm{u}), \notag\\
	& \Delta_{ij}   \stackrel{iid}{\sim} \text{Bernoulli}(p), \notag
\end{align}
where $p$ is the proportion of records that match, and $\bm{m}$ is the set of parameters for the matching record pairs $m_{fl} = P(\gamma_{ij}^f = l|\Delta_{ij} = 1)$ for all fields $f$ and agreement levels $l$, and $\bm{u}$ is defined similarly for non-matching pairs. These $\bm{m}$ and $\bm{u}$ parameters are then used to calculate the likelihood ratio  
\begin{align}
	\label{eqn:wts}
	w_{ij} = \frac{P(\gamma_{ij} \mid \Delta_{ij} = 1)}{P(\gamma_{ij} \mid \Delta_{ij} = 0)}.
\end{align}
This ratio will be large when there is strong evidence of the pair being a match, and small otherwise.

The Fellegi Sunter method then uses thresholds $T_m$ and $T_u$ such that all record pairs with $w_{ij} > T_m$ are declared as links, and all those with $w_{ij} < T_u$ are declared as non-links (with those such that $T_u < w_{ij} < T_m$ left undetermined and subject to clerical review). This is a problem in practice as transitive closures can be violated. Since each $w_{ij}$ is calculated independently, there can be situations in which both $w_{ij} > T_m$ and $w_{i'j} > T_m$, so both $(i,j)$ and $(i', j)$ are declared as links, even though such a matching is not allowed by assumption.

To address this issue, \cite{jaro1989} proposed solving the following linear sum assignment problem:
\begin{align}
	\label{eqn:jaro}
	&\max_{\Delta} \sum_{i=1}^{n_1} \sum_{j=1}^{n_2} w_{ij} \Delta_{ij} 
	\quad \text{subject to} \quad \Delta_{ij} \in \{0,1\}; \\
	& \quad \quad \sum_{i=1}^{n_1}  \Delta_{ij}  \leq 1, j=1,2, \ldots n_2; \; \text{and} \notag\\
	& \quad \quad\sum_{i=1}^{n_2}  \Delta_{ij}  \leq 1, i=1,2, \ldots n_1. \notag
\end{align}
